### 2021-10-27 manuscript/fig_build_time.tex
odgi build performance figure

\begin{figure}[h!]
	\centering
	\includegraphics[width=\linewidth]{fig/build/mean_build_time.pdf}
	\caption{Performance evaluation of \textit{odgi build} when translating a 90 haplotypes human chromosome 6 graph into ODGI's native format. The error margins are the standard deviations computed from 10 runs for each of the different threading configurations.}
	\label{fig:build}
\end{figure}