### 2021-11-01 manuscript/fig_sorting.tex

\begin{figure}[h!]
	\begin{subfigure}{\linewidth}
		%\caption{}
		\centering
		% include first image
		\includegraphics[width=1.0\linewidth, trim=-0cm 2cm 0 0cm]{fig/sorting/chr6_pan_fa_a2fb268_4030258_6a1ecc2_smooth_C4_bad_sorted}
		\label{fig:bad-sorting}
	\end{subfigure}
%	\includegraphics[width=\linewidth]{fig/metrics/chr4.pan.HTTex1.gfa.multiqc_odgi_stats.png}
	\caption{
		Visualization of a poorly sorted C4 pangenome graph.
		Same \textit{odgi viz} visualization as the one shown in Fig.~\ref{fig:odgi_viz}\textbf{e}, but with a graph where the bad node order hides the underlying copy number variation status present in the C4 region.
	}
	\label{fig:sorting}
\end{figure}