### 2021-11-04 manuscript/fig_build_time.tex
passing through

\begin{figure}[ht!]
%  \centering
\floatbox[{\capbeside\thisfloatsetup{capbesideposition={left,top},capbesidewidth=0.45\linewidth}}]{figure}[\FBwidth]
         {\caption{Performance evaluation of \textit{odgi build} when translating a 90-haplotype graph of human chromosome 6 into ODGI's native format.
             Build time decreases as parallelism increases, but with diminishing per-thread processing rates.
           }\label{fig:build}}
         {\includegraphics[width=\linewidth]{fig/build/mean_build_time.pdf}} %\includegraphics[width=5cm]{name}}
	     %\includegraphics[width=\linewidth]{fig/build/mean_build_time.pdf}
	     %\caption{Performance evaluation of \textit{odgi build} when translating a 90-haplotype graph of human chromosome 6 into ODGI's native format.}

\end{figure}