### 2022-02-14 manuscript/table_viz.tex

\begin{table}[!ht]
	\centering
	\caption{TODO: PLACE THE CAPTION ON THE TOP. CONDITIONALLY DIFFICULT LATEX.}
	\begin{tabular}{|l|l|l|l|l|}
		\hline
		\textbf{haps} & \textbf{time\_odgi\_viz} & \textbf{time\_vg\_viz} & \textbf{memory\_odgi\_viz} & \textbf{memory\_vg\_viz} \\ \hline \hline
		1 & \textbf{4.143} & 43.401* & \textbf{1.0401276} & 7.0868568* \\ \hline
		2 & \textbf{5.303} & 57.94** & \textbf{1.1399148} & 9.6889932** \\ \hline
		4 & \textbf{7.852} & 76.151** & \textbf{1.2726568} & 14.047634** \\ \hline
		8 & \textbf{12.951} & 109.16** & \textbf{1.5252528} & 21.839728** \\ \hline
		16 & \textbf{22.334} & 170.72** & \textbf{1.8844984} & 36.9709848** \\ \hline
		32 & \textbf{49.236} & 303.397** & \textbf{2.8689692} & 69.8805692** \\ \hline
		64 & \textbf{102.816} & 543.5** & \textbf{4.7780848} & 127.3594708** \\ \hline
	\end{tabular}
\end{table}